---
title: "BPM Code Walkthrough"
author: "Preston Stevenson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basketball Reference's BPM R Walk-through

There are a handful of public NBA one number metrics available to assess the value of a player. Most of these will provide similar player tiers but can be traced back to a handful of different modeling methods. The models are used to calculate the coefficients used to adjust players box scores and/or tracking stats. This walk-through does not include the modeling portion of BPM, but it can be a fun exercise to get a bit more comfortable with cleaning and combining data while also working with bigger functions.

BPM stands for Box Plus-Minus and it is [Basketball Reference's](https://www.basketball-reference.com/) one number metric that was created and developed by [Daniel Myers](https://twitter.com/DSMok1/). BPM uses only box score metrics, so there is no tracking data included in the calculation which will lead to a less robust model than something like [DARKO's DPM](https://apanalytics.shinyapps.io/DARKO//) and [FiveThirtyEights RAPTOR](https://projects.fivethirtyeight.com/nba-player-ratings/) There is a lot more information available for every step I am doing [here](https://www.basketball-reference.com/about/bpm2.html/) I used the attached Google sheet [here](https://docs.google.com/spreadsheets/d/1PhD9eo3IqzpQo21-yVJPQzYjpXl_h-ZonIKqGEKBqwY/edit#gid=307166562/) to follow the calculation.

## Code

```{r packages}
library(tidyverse)
library(nbastatR)
library(rvest)
library(janitor)
```

I separate the player and team stats pull from nbastatR for speed and to keep it from breaking. You may have to chunk it into a few pieces and rbind the data if it stops.

```{r nbastatR}
box_scores(game_ids = 22200001:22201230,
           box_score_types = c("Traditional", "Advanced"),
           result_types = c("player"), 
           join_data = TRUE,
           assign_to_environment = TRUE,
           return_message = TRUE)
p_stats22 <- dataBoxScorePlayerNBA

box_scores(game_ids = 22200001:22201230,
           box_score_types = c("Traditional", "Advanced"),
           result_types = c("team"), 
           join_data = TRUE,
           assign_to_environment = TRUE,
           return_message = TRUE)
t_stats22 <- dataBoxScoreTeamNBA
```

I remove game 193 because the advanced stats from [NBA.com](https://www.nba.com/stats) do not populate and the nulls will mess up the functions.

```{r}
#Games advanced stats come over as NA's
t_stats22 <- t_stats22 %>%
  filter(idGame != 22200193)
```

It is important to know everyone's positions for BPM, because different positions receiving certain stats are worth different values (an assist for a center is worth more than an assist for a guard), so we have to gather all players' positions from [Basketball Reference](https://www.basketball-reference.com/).

```{r}
#getting player positions from bbref play by play table
url <- "https://www.basketball-reference.com/leagues/NBA_2023_play-by-play.html"

df_pos <-  url %>%
  read_html() %>% 
  html_table()

df_pos <- df_pos[[1]]

#convert first row to header names and then clean up the variable name
df_pos <- df_pos %>% 
  row_to_names(row_number = 1) %>% 
  clean_names()

#table on website is broken up with aggregated rows, this removes those
df_pos <- df_pos %>% 
  group_by(player) %>% 
  mutate(team_count = n()) %>% 
  ungroup() %>% 
  mutate(keep = case_when(
    team_count == 1 | tm == "TOT" ~ "keep",
    TRUE ~ "Discard"
  )) %>% 
  filter(keep == "keep") 

df_pos <- df_pos %>% 
  select(player, ends_with("_percent")) %>% 
  pivot_longer(!player, names_to = "position", values_to = "time") %>% 
  mutate(time = parse_number(time)/100, 
         time = case_when(
           is.na(time) ~ 0, 
           TRUE ~ time))

#Find the position that each player spends the most time at, pivot wide
df_pos <- df_pos %>% 
  filter(player != "Player") %>% 
  group_by(player) %>% 
  arrange(desc(time)) %>% 
  mutate(rank = row_number(),
         pos = case_when(rank == 1 ~ position)) %>% 
  fill(pos) %>% 
  select(-rank) %>% 
  ungroup() %>%
  pivot_wider(names_from = position, values_from = time)

#Rename positions
df_pos <- df_pos %>% 
  mutate(pos = case_when(
    pos == "c_percent" ~ "C",
    pos == "pf_percent" ~ "PF",
    pos == "sf_percent" ~ "SF",
    pos == "sg_percent" ~ "SG",
    pos == "pg_percent" ~ "PG"
  ))
```

Next, we need to include all of the position values like a PG is a 1 and so on, so we can combine those to each player.

```{r}
#creating position value chart
pos <- c('PG','SG','SF','PF','C')
pos_val <- c(1, 2, 3, 4, 5)
pos_value <- as.data.frame(cbind(pos,pos_val))

#joining positions and position values
play_pos22 <- merge(x=df_pos, y=pos_value, by.x = "pos", by.y = "pos", all.x = TRUE) 
play_pos22$pos_val <- as.numeric(play_pos22$pos_val)
play_pos22 <- play_pos22 %>%
  select(player, pos, pos_val)
```

We need to use the nbastatR function again to pull a table that connects a players BBrefID to their NBA ID. This is important because not all players' names match identically from website to website, primarily overseas players with non-English letters, or maybe you drafted two people with the same name the same year...

```{r}
#This is helpful for a lot of website scraping related activities.
Sys.setenv("VROOM_CONNECTION_SIZE"= 131072 * 10)
#gathering player ids to join with NBA IDs
bref_players_stats(seasons = 2023, tables=c("totals"))
bref_ids <- dataBREFPlayerTotals %>%
  select(namePlayer, slugPlayerBREF, idPlayerNBA)
bref_ids[is.na(bref_ids$idPlayerNBA),]

#brefPlayers <- dictionary_bref_players()

#there are some ID's that the package did not return, we will have to add these.
#their id is the number in the link on nba.com
bref_ids <- bref_ids %>% 
  mutate(idPlayerNBA = case_when(
    namePlayer == "Xavier Tillman Sr." ~ 1630214, 
    namePlayer == "OG Anunoby" ~ 1628384, 
    namePlayer == "John Butler" ~ 1631219, 
    namePlayer == "Jeff Dowtin" ~ 1630288, 
    namePlayer == "A.J. Green" ~ 1631260, 
    TRUE ~ idPlayerNBA
  ))
```

nbastatR's dictionary_bref_players() function is throwing an error not allowing me to have the table that I mentioned above. This will affect a little under 5% of the players, and I will default their position at 3 until this error is resolved. There are position adjustments taking into account a players average stats that will skew their starting position towards their real value, so the damage for the final product is minimal. So, the bref_ids section is useless right now.

```{r}
p_stats22 <- merge(x=p_stats22, y=play_pos22, by.x = "namePlayer", by.y = "player", all.x = TRUE)
p_stats22$pos_val[is.na(p_stats222$pos_val)] <- 3
p_stats22$pos_val <- as.numeric(p_stats222$pos_val)
```

I am not going to go into much detail on the reasoning behind these coefficients. Again, look [here](https://www.basketball-reference.com/about/bpm2.html/) for a full explanation. They vary linearly and adjust for the difference in certain statistic performance from a big to a guard like mentioned before. Offensive rebounds are better for guards, while assists are worth more for a big and so forth.

```{r}
#Create table for BPM Coefficients
pos1 <- c(0.86,	-0.56,	-0.2464,	0.389,	0.58,	-0.964,	0.613,	0.116,	0.0,	1.369,	1.327,	-0.367)
pos5 <- c(0.86,	-0.78,	-0.3432,	0.389,	1.034,	-0.964,	0.181,	0.181,	0.0,	1.008,	0.703,	-0.367)
co_bpm <- as.data.frame(rbind(pos1, pos5))
colnames(co_bpm) <- c("Adj. Pt",	"FGA",	"FTA",	"FG3Pt(bonus)",	"AST",	"TO",
                      "ORB", "DRB", "TRB", "STL", "BLK", "PF")

#create table for OBPM Coefficients
opos1 <- c(0.605,	-0.33, -0.1452, 0.477, 0.476, -0.579,	0.606, -0.112, 0.0, 0.177, 0.725, -0.439)
opos5 <- c(0.605, -0.472, -0.20768, 0.477, 0.476, -0.882, 0.422, 0.103, 0.0, 0.294, 0.097, -0.439)
co_obpm <- as.data.frame(rbind(opos1, opos5))
colnames(co_obpm) <- c("Adj. Pt",	"FGA",	"FTA",	"FG3Pt(bonus)",	"AST",	"TO",
                       "ORB", "DRB", "TRB", "STL", "BLK", "PF")

#create table for positions percent for overall and offensive roles
positions <- c(2.130, 8.668, -2.486, 0.992, -3.536, 1.667, 50)
num <- c(1:7)
pos_per <- as.data.frame(rbind(positions,num))
pos_per <- pos_per[1,]
colnames(pos_per) <- c("intercept", "pct_TRB", "pct_STL", "pct_PF", "pct_AST", "pct_BLK", "Min_Wt")

offrole <- c(6.000, -6.642, -8.544, -0.330, 4.000, 50)
num2 <- c(1:6)
off_role_pos <- as.data.frame(rbind(offrole,num2))
off_role_pos <- off_role_pos[1,]
colnames(off_role_pos) <- c("intercept", "pct_ast", "pct_thresh_points", "pt_thresh", "Default_pos", "Min_Wt")

#This helps to iterate over all teams later
teamlist <- c('MIA', 'ORL', 'ATL', 'CHA', 'WAS',
              'BKN', 'NYK', 'PHI', 'TOR', 'BOS',
              'MIL', 'CHI', 'CLE', 'IND', 'DET',
              'LAL', 'LAC', 'SAC', 'GSW', 'PHX',
              'DAL', 'SAS', 'NOP', 'HOU', 'MEM',
              'UTA', 'DEN', 'POR', 'OKC', 'MIN')

```

Now it is time for the main course. There are a few functions that basically fill out everything needed to calculate BPM, and by the end, it is over 130 fields. The core groups needed are player box score statistics per 100 possessions, team adjusted efficiency statistics per 100 possessions, and adjusted player positions. The first function contains the bulk of the numbers needed. I go back and forth between calculating team and player stats, so I could follow along with Daniel's table.

```{r}
bpm_tab <- function(team) {
  #filter for team and then get totals for min, pts, fga, etc
  d1 <- t_stats22 %>%
    filter(team == slugTeam) %>%
    mutate(G = 1) %>%
    summarize(minExact = sum(minExact), pts = sum(pts),
              fga = sum(fga), fta = sum(fta), pace = mean(pace), 
              ortg = mean(ortg), drtg = mean(drtg), netrtg = mean(netrtg),
              treb = sum(treb), stl = sum(stl), pf = sum(pf), blk = sum(blk),
              ast = sum(ast), ortg = mean(ortg), drtg = mean(drtg))
  #add tm points/tsa and then baseline points
  d1 <- d1 %>%
    mutate(tm_points_tsa = pts/(fga + (fta*(-.2464/-.56))),
           baseline_pts = 1) 
  # new dataset to look at individual players
  #filter by team, add games so it will count total games per player
  #group by player and team then sum up all the stats excluding percentages
  d2 <- p_stats222 %>%
    filter(team == slugTeam) %>%
    mutate(G = 1) %>%
    group_by(namePlayer, slugTeam) %>%
    summarize(G = sum(G), pts = sum(pts), ast = sum(ast), stl = sum(stl), blk = sum(blk),
              plusminus = sum(plusminus), pctFG = mean(pctFG), pct3FG = mean(pctFG3), MP = sum(minExact),
              fga = sum(fga), fta = sum(fta), fgm = sum(fgm), ftm = sum(ftm), pctFT = mean(pctFT), fg3m = sum(fg3m),
              fg2a = sum(fg2a), fg2m = sum(fg2m), pctFG2 = mean(pctFG2), dreb = sum(dreb), oreb = sum(oreb),
              treb = sum(treb), tov = sum(tov), pf = sum(pf), pos_val = mean(pos_val))
  #start adding advanced stats, some fields use other mutated fields, it is broken up because R gets confused sometimes
  d2 <- d2 %>%
    mutate(efg = (fgm + .5*fg3m)/fga,
           tsa = fga + (.44*fta),
           pts_tsa = pts/tsa)
  #same case here
  d2 <- d2 %>%
    mutate(adj_pts = (pts_tsa-d1$tm_points_tsa+1)*tsa, 
           poss = (MP*d1$pace)/48)
  d2$adj_pts[is.na(d2$adj_pts)] <- 0
  #same thing, adding all the per 100 stats now
  d2 <- d2 %>%
    mutate(adj_pts100 = adj_pts/poss*100, 
           fga100 = fga/poss*100,
           fta100 = fta/poss*100,
           fg3m100 = fg3m/poss*100,
           ast100 = ast/poss*100,
           tov100 = tov/poss*100,
           oreb100 = oreb/poss*100,
           dreb100 = dreb/poss*100,
           treb100 = treb/poss*100,
           stl100 = stl/poss*100,
           blk100 = blk/poss*100,
           pf100 = pf/poss*100,
           thresh_pts = tsa * (pts_tsa-(d1$tm_points_tsa+off_role_pos$pt_thresh)),
           pct_MP = MP/(d1$minExact/5))
  d2$thresh_pts[is.na(d2$thresh_pts)] <- 0
  
  #after adding threshold points to d2 we need team threshold points in d1
  d1 <- d1 %>%
    mutate(tm_thresh_pts = sum(d2$thresh_pts, na.rm = TRUE))
  
  #adding pct of stats for when player is on the floor
  d2 <- d2 %>%
    mutate(pct_stl = stl/d1$stl/pct_MP,
           pct_treb = treb/d1$treb/pct_MP,
           pct_blk = blk/d1$blk/pct_MP,
           pct_ast = ast/d1$ast/pct_MP,
           pct_pf = pf/d1$pf/pct_MP,
           pct_thresh_pts = thresh_pts/d1$tm_thresh_pts/pct_MP)
  
  #calculating estimated position using average position stat table against their pct stats
  d2 <- d2 %>%
    mutate(est_pos_1 = pos_per[1,1] + (pct_treb*pos_per$pct_TRB) + (pct_stl*pos_per$pct_STL)
           + (pct_blk*pos_per$pct_BLK) + (pct_ast*pos_per$pct_AST)
           + (pct_pf*pos_per$pct_PF))
  #starting section of adding all of the adjusted positions until the team avg rounds to 3
  #round 1
  d2 <- d2 %>%
    mutate(min_adj_1 = ((est_pos_1*MP) + (pos_val*pos_per[1,7]))/ (MP+pos_per[1,7]))
  d2 <- d2 %>%
    mutate(trim_1 = max(min(min_adj_1, 5), 1))
  d2 <- d2 %>%
    mutate(tm_avg_top1 = trim_1*MP)
  d1 <- d1 %>%
    mutate(tm_avg_1 = sum(d2$tm_avg_top1)/minExact)
  #round 2 
  d2 <- d2 %>%
    mutate(adj_pos_2 = min_adj_1 - (d1$tm_avg_1 - 3))
  d2 <- d2 %>%
    mutate(trim_2 = max(min(adj_pos_2, 5), 1))
  d2 <- d2 %>%
    mutate(tm_avg_top2 = trim_2*MP)
  d1 <- d1 %>%
    mutate(tm_avg_2 = sum(d2$tm_avg_top2)/minExact)
  #round 3
  d2 <- d2 %>%
    mutate(adj_pos_3 = min_adj_1 - (d1$tm_avg_1 - 3) - (d1$tm_avg_2 - 3))
  d2 <- d2 %>%
    mutate(trim_3 = max(min(adj_pos_3, 5), 1))
  d2 <- d2 %>%
    mutate(tm_avg_top3 = trim_3*MP)
  d1 <- d1 %>%
    mutate(tm_avg_3 = sum(d2$tm_avg_top3)/minExact)
  #round 4
  d2 <- d2 %>%
    mutate(adj_pos_4 = min_adj_1 - (d1$tm_avg_1 - 3) - (d1$tm_avg_2 - 3) - (d1$tm_avg_3 - 3))
  d2 <- d2 %>%
    mutate(trim_4 = max(min(adj_pos_4, 5), 1))
  d2 <- d2 %>%
    mutate(tm_avg_top4 = trim_4*MP)
  d1 <- d1 %>%
    mutate(tm_avg_4 = sum(d2$tm_avg_top4)/minExact)
  #finishing and calculating on the round 5
  d2 <- d2 %>%
    mutate(adj_pos_5 = min_adj_1 - (d1$tm_avg_1 - 3) - (d1$tm_avg_2 - 3) 
           - (d1$tm_avg_3 - 3) - (d1$tm_avg_4 - 3))
  d2 <- d2 %>%
    mutate(position_adj = max(min(adj_pos_5, 5), 1))
  
  #starting similar process but calculating offensive position value
  d2 <- d2 %>%
    mutate(est_off_role = off_role_pos$intercept + (off_role_pos$pct_ast*pct_ast) 
           + (off_role_pos$pct_thresh_points*pct_thresh_pts))
  d2$est_off_role[is.na(d2$est_off_role)] <- 0
  d2 <- d2 %>%
    mutate(o_min_adj_1 = ((est_off_role*MP) + (off_role_pos$Default_pos*off_role_pos$Min_Wt))/
             (MP + off_role_pos$Min_Wt))
  d2 <- d2 %>%
    mutate(o_trim_1 = max(min(o_min_adj_1, 5), 1))
  d2 <- d2 %>%
    mutate(o_tm_avg_top1 = o_min_adj_1*MP)
  #removing NA's
  d2$o_tm_avg_top1[is.na(d2$o_tm_avg_top1)] <- 0
  d1 <- d1 %>%
    mutate(o_tm_avg_1 = sum(d2$o_tm_avg_top1)/minExact)
  #round 2
  d2 <- d2 %>%
    mutate(o_adj_2 = (o_min_adj_1 - (d1$o_tm_avg_1 - 3)))
  d2 <- d2 %>%
    mutate(o_trim_2 = max(min(o_adj_2, 5), 1))
  d2 <- d2 %>%
    mutate(o_tm_avg_top2 = o_adj_2*MP)
  d2$o_tm_avg_top2[is.na(d2$o_tm_avg_top2)] <- 0
  d1 <- d1 %>%
    mutate(o_tm_avg_2 = sum(d2$o_tm_avg_top2)/minExact)
  #round 3
  d2 <- d2 %>%
    mutate(o_adj_3 = (o_min_adj_1 - (d1$o_tm_avg_1 - 3) - (d1$o_tm_avg_2 - 3)))
  d2 <- d2 %>%
    mutate(o_trim_3 = max(min(o_adj_3, 5), 1))
  d2 <- d2 %>%
    mutate(o_tm_avg_top3 = o_adj_3*MP)
  d2$o_tm_avg_top3[is.na(d2$o_tm_avg_top3)] <- 0
  d1 <- d1 %>%
    mutate(o_tm_avg_3 = sum(d2$o_tm_avg_top3)/minExact)
  #round 4 is used as the official numbers, once the tm avg gets to 3 for all teams
  d2 <- d2 %>%
    mutate(o_adj_4 = (o_min_adj_1 - (d1$o_tm_avg_1 - 3) - (d1$o_tm_avg_2 - 3) - (d1$o_tm_avg_3 - 3)))
  d2 <- d2 %>%
    mutate(o_position_adj = max(min(o_adj_4, 5), 1))
  #BPM coefficients to use based on positions
  d2 <- d2 %>%
    mutate(co_adj_pts = ((5 - position_adj)/4*co_bpm[1,1]) 
           + ((position_adj - 1)/4*co_bpm[2,1])) %>%
    mutate(co_fga = ((5 - o_position_adj)/4*co_bpm[1,2]) 
           + ((o_position_adj - 1)/4*co_bpm[2,2])) %>%
    mutate(co_fta = ((5 - o_position_adj)/4*co_bpm[1,3]) 
           + ((o_position_adj - 1)/4*co_bpm[2,3])) %>%
    mutate(co_fg3m = ((5 - position_adj)/4*co_bpm[1,4]) 
           + ((position_adj - 1)/4*co_bpm[2,4])) %>%
    mutate(co_ast = ((5 - position_adj)/4*co_bpm[1,5]) 
           + ((position_adj - 1)/4*co_bpm[2,5])) %>%
    mutate(co_tov = ((5 - position_adj)/4*co_bpm[1,6]) 
           + ((position_adj - 1)/4*co_bpm[2,6])) %>%
    mutate(co_oreb = ((5 - position_adj)/4*co_bpm[1,7]) 
           + ((position_adj - 1)/4*co_bpm[2,7])) %>%
    mutate(co_dreb = ((5 - position_adj)/4*co_bpm[1,8]) 
           + ((position_adj - 1)/4*co_bpm[2,8])) %>%
    mutate(co_treb = ((5 - position_adj)/4*co_bpm[1,9]) 
           + ((position_adj - 1)/4*co_bpm[2,9])) %>%
    mutate(co_stl = ((5 - position_adj)/4*co_bpm[1,10]) 
           + ((position_adj - 1)/4*co_bpm[2,10])) %>%
    mutate(co_blk = ((5 - position_adj)/4*co_bpm[1,11]) 
           + ((position_adj - 1)/4*co_bpm[2,11])) %>%
    mutate(co_pf = ((5 - position_adj)/4*co_bpm[1,12]) 
           + ((position_adj - 1)/4*co_bpm[2,12]))
  
  #raw bpm calculation for different categories
  d2 <- d2 %>%
    mutate(scoring = (co_adj_pts*adj_pts100) + (co_fga*fga100) 
           + (co_fta*fta100) + (co_fg3m*fg3m100)) %>%
    mutate(ballhandling = (co_ast*ast100) + (co_tov*tov100)) %>%
    mutate(rebounding = (co_oreb*oreb100) + (co_dreb*dreb100) + (co_treb*treb100)) %>%
    mutate(defense = (co_stl*stl100) + (co_blk*blk100) + (co_pf*pf100))
  
  d2 <- d2 %>%
    mutate(pos_constant = if(position_adj < 3){
      (3 - position_adj)/2*-.818
    } else{
      1.387*(o_position_adj-3)
    }) 
  d2 <- d2 %>%
    mutate(rawbpm = scoring + ballhandling + rebounding + defense + pos_constant)
  d2 <- d2 %>%
    mutate(o_co_adj_pts = ((5 - position_adj)/4*co_obpm[1,1]) 
           + ((position_adj - 1)/4*co_obpm[2,1])) %>%
    mutate(o_co_fga = ((5 - o_position_adj)/4*co_obpm[1,2]) 
           + ((o_position_adj - 1)/4*co_obpm[2,2])) %>%
    mutate(o_co_fta = ((5 - o_position_adj)/4*co_obpm[1,3]) 
           + ((o_position_adj - 1)/4*co_obpm[2,3])) %>%
    mutate(o_co_fg3m = ((5 - position_adj)/4*co_obpm[1,4]) 
           + ((position_adj - 1)/4*co_obpm[2,4])) %>%
    mutate(o_co_ast = ((5 - position_adj)/4*co_obpm[1,5]) 
           + ((position_adj - 1)/4*co_obpm[2,5])) %>%
    mutate(o_co_tov = ((5 - position_adj)/4*co_obpm[1,6]) 
           + ((position_adj - 1)/4*co_obpm[2,6])) %>%
    mutate(o_co_oreb = ((5 - position_adj)/4*co_obpm[1,7]) 
           + ((position_adj - 1)/4*co_obpm[2,7])) %>%
    mutate(o_co_dreb = ((5 - position_adj)/4*co_obpm[1,8]) 
           + ((position_adj - 1)/4*co_obpm[2,8])) %>%
    mutate(o_co_treb = ((5 - position_adj)/4*co_obpm[1,9]) 
           + ((position_adj - 1)/4*co_obpm[2,9])) %>%
    mutate(o_co_stl = ((5 - position_adj)/4*co_obpm[1,10]) 
           + ((position_adj - 1)/4*co_obpm[2,10])) %>%
    mutate(o_co_blk = ((5 - position_adj)/4*co_obpm[1,11]) 
           + ((position_adj - 1)/4*co_obpm[2,11])) %>%
    mutate(o_co_pf = ((5 - position_adj)/4*co_obpm[1,12]) 
           + ((position_adj - 1)/4*co_obpm[2,12]))
  d2 <- d2 %>%
    mutate(o_scoring = (o_co_adj_pts*adj_pts100) + (o_co_fga*fga100) 
           + (o_co_fta*fta100) + (o_co_fg3m*fg3m100)) %>%
    mutate(o_ballhandling = (o_co_ast*ast100) + (o_co_tov*tov100)) %>%
    mutate(o_rebounding = (o_co_oreb*oreb100) + (o_co_dreb*dreb100) + (o_co_treb*treb100)) %>%
    mutate(o_defense = (o_co_stl*stl100) + (o_co_blk*blk100) + (o_co_pf*pf100))
  d2 <- d2 %>%
    mutate(o_pos_constant = if(position_adj < 3){
      (3 - position_adj)/2*-1.698
    } else{
      .43*(o_position_adj-3)
    }) 
  d2 <- d2 %>%
    mutate(rawobpm = o_scoring + o_ballhandling + o_rebounding + o_defense + o_pos_constant)
  d2 <- d2 %>%
    mutate(rawbpm_cont = rawbpm*pct_MP) %>%
    mutate(rawobpm_cont = rawobpm*pct_MP) %>%
    mutate(MPG = MP/G) %>%
    mutate(reMPG = MP/(G+4)) %>%
    mutate(reMin = max((450-MP)/3, 0)) %>%
    mutate(expBPM = 4.75 + .175*reMPG)
  
  print(d2)
}

```

We are left with raw BPM that does not properly account for the context of their entire team's talent level. We gather offensive rating, defensive rating, pace, and calculate average team lead(This is important because it has been proven over and over that players with the lead play worse, so just summing up to the teams efficiency could be misleading). The team's adjusted efficiency is subtracted from the players' raw BPM to land on a final BPM. There are some extras that can be calculated from BPM like VORP and reBPM. Finally there is a function that runs this over every team and provides BPM, OBPM, DBPM, etc.

```{r}
teamrating <- function(team){
  
  d0 <- bpm_tab(team)
  #This is used to calculate the NBA's overall ortg
  d1 <- t_stats22 %>%
    summarize(rtg = mean(ortg))
  
  d2 <- t_stats22 %>%
    filter(slugTeam == team) %>%
    mutate(G = 1) %>%
    summarize(tortg = mean(ortg), tdrtg = mean(drtg), pace = mean(pace), G = sum(G))
  
  d2 <- d2 %>%
    mutate(ortg = tortg-d1$rtg) %>%
    mutate(drtg = d1$rtg-tdrtg) %>%
    select(ortg, drtg, pace, G) %>%
    mutate(Team_rtg = ortg + drtg) %>%
    mutate(avg_lead = Team_rtg*pace/100/2) %>%
    mutate(lead_bonus = .35/2*avg_lead) %>%
    mutate(adj_tm_rtg = Team_rtg + lead_bonus) %>%
    mutate(adj_o_rtg = ortg + lead_bonus/2)
  d3 <- d0 %>%
    select(namePlayer, rawbpm_cont, rawobpm_cont)
  d3 <- d3[,2:3]
  d3 <- d3 %>%
    summarize(rawbpm_cont = sum(rawbpm_cont), rawobpm_cont = sum(rawobpm_cont))
  d3 <- d3 %>%
    mutate(adj_rating = (d2$adj_tm_rtg - rawbpm_cont)/5) %>%
    mutate(adj_orating = (d2$adj_o_rtg - rawobpm_cont)/5)
  d2 <- cbind(d2,d3)
  print(d2)
}

bpm <- function(team) {
  
  d1 <- bpm_tab(team)
  d2 <- teamrating(team)
  d1 <- d1 %>%
    mutate(BPM = rawbpm + d2$adj_rating) %>%
    mutate(OBPM = rawobpm + d2$adj_orating) %>%
    mutate(DBPM = BPM - OBPM) %>%
    mutate(Contribution = BPM*pct_MP) %>%
    mutate(VORP = (BPM + 2) * pct_MP*d2$G/82) %>%
    mutate(reBPM = (BPM*MP + reMin*expBPM)/(MP+reMin)) %>%
    mutate(reOBPM = (OBPM*MP + reMin*expBPM)/(MP+reMin)) %>%
    mutate(reDBPM = reBPM - reOBPM) %>%
    select(namePlayer, slugTeam, position_adj, o_position_adj, MP, MPG, BPM, OBPM, DBPM,       Contribution, VORP, reMPG, reMin, expBPM, reBPM, reOBPM, reDBPM) %>%
    arrange(desc(BPM))
  print(d1)
}

all_bpm <- function(){
  output = data.frame()
  for (i in teamlist) {
    
    result <- bpm(i)
    output <- rbind(output, result)
  }
  print(output)
}

bpm22 <- all_bpm()
```

This has been a fun experience to extrapolate all of Daniel Myers' calculations and turn it into a fairly quick running script outside of scraping the data. I know that there are probably plenty of ways to improve this codes readability and efficiency, but hopefully it has been helpful to someone and it shows a little bit of what I can do and enjoy.
